---
title: "Clean PPCS 1996"
author: "Brenda Fried"
date: "7/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries

```{r load-libraries}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, asciiSetupReader, here)
```

#Load Data
```{r load-data}
ppcs_1996 <- read_ascii_setup(here('raw_data', '06999-0001-Data.txt'),here('raw_data', '06999-0001-Setup.sas'))

```

#Civilian Race
```{r civilian-race}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_race = case_when(
    (RESPONDENT_RACE == "White" & RESPONDENT_ETHNICITY == "Non-Hispanic") ~ "white",
    (RESPONDENT_RACE == "Black" ) ~ "black",
    (RESPONDENT_ETHNICITY == "Hispanic" & RESPONDENT_RACE != "Black") ~ "hispanic",
    TRUE ~ "other"
  ))
```

#Age
```{r civilian-age}
#Age - same as RESPONDENT_AGE column
ppcs_1996 <- ppcs_1996 %>% 
  mutate(civilian_age = NCVS_RESPONDENT_AGE)

```
#Gender
```{r civilian-gender}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_gender = case_when(
    (RESPONDENT_SEX == "Male") ~ 1,
    (RESPONDENT_SEX == "Female") ~ 2
))

```

#Civilian income

```{r civilian-income}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_income = NA)

```

#civilian employed
not found so coded as 0
```{r civilian-employed}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_employed = 0)
```
#Population size

```{r pop-size}
ppcs_1996 <- ppcs_1996 %>%
  mutate(population_size = NA)
```
#Time of Encounter
```{r time-of-encounter}
ppcs_1996 <- ppcs_1996 %>%
  mutate(time_of_encounter = case_when(
    (DAYTIME_OR_NIGHTTIME_WHEN_INCIDENT == "Daytime") ~ 3,
    (DAYTIME_OR_NIGHTTIME_WHEN_INCIDENT == "Nighttime") ~ 6
))
```
#Officer race

```{r officer-race-black}
#officer_race
ppcs_1996 <- ppcs_1996 %>%
  mutate(off_black = case_when(
    (OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'All black' | OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'Mostly black' | OFFICER_WHITE_BLACK_OR_OTHER_RACE == 'Black' ) ~ 1,
    TRUE ~ 0
  ))

```


```{r officer-race-white}
ppcs_1996 <- ppcs_1996%>%
  mutate(off_white = case_when(
    (OFFICERS_WHITE_BLACK_OR_OTHER_RACE == "All white" | OFFICERS_WHITE_BLACK_OR_OTHER_RACE == "Mostly white" | OFFICER_WHITE_BLACK_OR_OTHER_RACE == "White" ) ~ 1,
    TRUE ~ 0
  ))

```

```{r officer-race-other}
ppcs_1996 <- ppcs_1996 %>%
  mutate(off_other = case_when(
    (OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'All of some other race' | OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'Mostly some other race' | OFFICER_WHITE_BLACK_OR_OTHER_RACE == 'Other' ) ~ 1,
    TRUE ~ 0
  ))
```


```{r Officer-race-split}
ppcs_1996 <- ppcs_1996 %>%
  mutate(off_split = case_when(
    (OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'Equally mixed') ~1,
    TRUE ~ 0
  ))
```

```{r officer-race-hispanic}
ppcs_1996 <- ppcs_1996 %>%
  mutate(off_hispanic = 0)

```

#Type of incident
```{r type-of-incident}
happened <- c('Once', 'More than once')

ppcs_1996 <- ppcs_1996 %>%
  mutate(type_of_incident = case_when(
    (REASON_TRAFFIC_PARKING_VIOLATION %in% happened ) ~2,
    (REASON_INVOLVED_IN_TRAFFIC_ACCIDENT %in% happened)~3,
    (REASON_WITNESS_TO_TRAFFIC_ACCIDENT %in% happened)~3,
    (REASON_REPORT_A_CRIME %in% happened) ~ 3,
    (REASON_CRIME_THOUGHT_INVOLVED_IN %in% happened) ~ 3,
    (REASON_COMMUNITY_MEETINGS %in% happened) ~3,
    TRUE ~ NA_real_
  ))
```


#Civilian Behavior
```{r civilian-behavior}

columns <- c(colnames(ppcs_1996)[95:98], colnames(ppcs_1996)[100:103])

columns
is_not_missing <- function(x) {
  ifelse(x == "Not at all" | x == 'missing/ni/ou',0, 1)
}

is_not_missing <- Vectorize(is_not_missing)

ppcs_1996_tmp <- ppcs_1996 %>%
  mutate_at(vars(columns), is_not_missing) %>%
  select(columns)

ppcs_1996 <- ppcs_1996 %>%
  mutate(row_sum = rowSums(ppcs_1996_tmp)) %>%
  mutate(civilian_behavior = case_when(
    (row_sum > 0) ~ 1,
    TRUE ~ 0
  ))

```
#Alternative Outcomes
Civilian searched
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_searched = case_when(
    (DOING_IN_THE_AREA_FRISK_PAT == 'Yes' | CRIME_THOUGHT_INVOLVED_IN_FRISK_PAT == 'Yes') ~ 1,
    TRUE ~ 0
  ))
```
Civilian arrested
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_arrested = case_when(
    (REASON_HAD_WARRANT_FOR_YOUR_ARREST == 'Yes' | HANDCUFF_HAD_WARRANT_FOR_YOUR_ARREST == 'Yes') ~ 1,
    TRUE ~ 0
  )) 
```
Civilian guilty of illegal dugs, alcohol, or weapon
```{r}

ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_guilty_of_illegal = case_when(
    (OTHER_POSSESSION_OF_DRUGS == 'Yes' | OTHER_POSSESSION_OF_FIREARM_WEAPON == 'Yes') ~ 1,
    TRUE ~ 0
  ))
```
#Civilian Injured
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_injured = ifelse(WERE_YOU_INJURED_AS_A_RESULT_OF_INCIDENT == 'Yes', 1, 0))
```
#Excessive force
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(excess_force = ifelse(TAKE_ANY_FORMAL_OR_INFORMAL_ACTIONS == 'Yes', 1, 0))
```

#force
```{r force}
ppcs_1996 <- ppcs_1996 %>%
  mutate(any_force = case_when(
    (OFFICER_USE_ANY_FORM_OF_PHYSICAL_FORCE == 'Yes') ~ 1,
    (POLICE_DO_USE_SOME_OTHER_FORM_OF_FORCE == 'Yes') ~ 1,
    (POLICE_DO_USE_A_CHOKEHOLD == 'Yes') ~ 1,
    TRUE ~ 0
  )) %>%
  mutate(grab_push = case_when(
    (POLICE_DO_HOLD_YOU_BY_THE_ARM == 'Yes') ~ 1,
    (POLICE_DO_PUSH_YOU == 'Yes') ~ 1,
    TRUE ~ 0
  )) %>%
  mutate(hit_kick = case_when(
    (POLICE_DO_HIT_OR_PUNCH_YOU == 'Yes') ~ 1,
    (POLICE_DO_KICK_YOU == 'Yes') ~ 1,
    TRUE ~ 0
  )) %>%
  mutate(point_gun = case_when(
    (WEAPONS_FIREARM == 'Yes') ~ 1,
    TRUE ~ 0
  )) %>%
  mutate(handcuffed = case_when(
    (DID_A_POLICE_OFFICER_HANDCUFF_YOU == 'Yes') ~ 1,
    TRUE ~ 0
  )) %>%
  mutate(pepper_stun = case_when(
    (WEAPONS_CHEMICAL_OR_PEPPER_SPRAY == 'Yes') ~ 1,
    TRUE ~ 0
  ))
```


##Cleaned Data Frame
```{r add-columns-for-future-columns}
ppcs_1996 <- ppcs_1996 %>% 
  mutate(contact = LAST_12_MONTHS_CONTACT_W_POLICE_OFFICER, face_to_face = CONTACTS_W_POLICE_OFFICER_IN_PERSON, num_face_to_face = NA)
```

```{r cleaned-data}
ppcs_1996 <- ppcs_1996 %>%
  select(contact, 
         face_to_face, 
         num_face_to_face, 
         civilian_race, 
         civilian_age, 
         civilian_gender, 
         civilian_income, 
         civilian_employed, 
         population_size, 
         time_of_encounter, 
         off_black, 
         off_white, 
         off_other, 
         off_split, 
         off_hispanic, 
         type_of_incident, 
         civilian_behavior, 
         civilian_searched, 
         civilian_arrested, 
         civilian_guilty_of_illegal, 
         civilian_injured, 
         excess_force, 
         any_force, 
         grab_push, 
         hit_kick, 
         point_gun, 
         handcuffed, 
         pepper_stun
         )
```
#Add year column
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(year = 1996)
```

#Save Data
```{r}
save(ppcs_1996, file='clean_data/ppcs_1996.RData')
sessionInfo()
```

