---
title: "PPCS Graphs"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data}
library(tidyverse)
library(scales)
library(varhandle)

load("merged_ppcs.RData")
graph_ppcs <- merged_ppcs

```


```{r income bar graph}
# force graph comparing income and race


graph_ppcs$civilian_income <- recode_factor(graph_ppcs$civilian_income,
                                            "1" = '< $20,000',
                                            "2" = '$20,000 - $50,000',
                                            "3" = '> $50,000'
                                            )

graph_ppcs %>%
  filter(!is.na(civilian_income)) %>%
  group_by(civilian_income) %>%
  filter(force == 1) %>%
  ggplot() +
  geom_bar(aes(x = civilian_income, fill = civilian_race), color = "white", position = "dodge") +
  coord_flip() +
  xlab("Civilian Income") + 
  ylab("Number of Interactions Involving Force") +
  labs(fill = "Civilian Race") 

graph_ppcs %>%
  filter(civilian_race != "white") %>% 
  filter(!is.na(civilian_income)) %>%
  group_by(civilian_income) %>%
  filter(force == 1) %>%
  ggplot() +
  geom_bar(aes(x = civilian_income, fill = civilian_race), color = "white", position = "dodge") +
  coord_flip() +
  xlab("Civilian Income") + 
  ylab("Number of Interactions Involving Force") +
  labs(fill = "Civilian Race")

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
