---
title: "PPCS Graphs"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data}
library(tidyverse)
library(scales)
library(varhandle)
library(ggplot2)
library(modelr)
library(dplyr)
library(ggthemes) 

load("merged_ppcs.RData")
graph_ppcs <- merged_ppcs

```




```{r income bar graph}
# force graph comparing income and race
# normalize based on count of incidents involving force over number of incidents

graph_ppcs$civilian_income <- recode_factor(graph_ppcs$civilian_income,
                                            "1" = '< $20,000',
                                            "2" = '$20,000 - $50,000',
                                            "3" = '> $50,000'
                                            )

# graph_ppcs %>%
#   filter(!is.na(civilian_income)) %>%
#   group_by(civilian_income) %>%
#   filter(force == 1) %>%
#   ggplot() +
#   geom_bar(aes(x = civilian_income, fill = civilian_race), color = "white", position = "dodge") +
#   coord_flip() +
#   xlab("Civilian Income") + 
#   ylab("Number of Interactions Involving Force") +
#   labs(fill = "Civilian Race") 

graph_ppcs %>%
  filter(!is.na(civilian_income)) %>%
  group_by(civilian_income, civilian_race) %>%
  mutate(incident_count = n()) %>%
  filter(force == 1) %>%
  summarize(force_count = n(), incident_count = mean(incident_count)) %>%
  mutate(prop = force_count / incident_count) %>%
  ggplot() +
  geom_col(aes(x = civilian_income, y = prop, fill = civilian_race), color = "white", position = "dodge") +
  coord_flip() +
  xlab("Civilian Income") + 
  ylab("Number of Interactions Involving Force") +
  labs(fill = "Civilian Race") 

```


```{r polar-and-flip-plot-of-force-by-race}

x <- merged_ppcs %>% filter(civilian_race != "white") %>% group_by(civilian_race) %>% summarize(race_count = n())

plot <- merged_ppcs %>%
  filter(civilian_race != "white") %>%
  group_by(civilian_race, force) %>%
  filter(force == 1) %>%
  summarize(force_count = n())

plot_ <- full_join(plot, x)

plot_ <-plot_ %>%  
  mutate(prop = force_count / race_count) 


bar <-  ggplot(data = plot_,  aes(x=civilian_race, y=prop, fill=force_count)) +
  geom_bar( stat="identity", position = "dodge") + 
  labs(fill = "Size of civilian race that force was used") +
  xlab('Civilian Race') +
  ylab('Proportion of force')


bar
bar + coord_polar()
bar + coord_flip()
```


```{r male and female graphs}
# force graph comparing gender
graph_ppcs$civilian_gender <- recode_factor(graph_ppcs$civilian_gender,
                                            "male" = 'Male',
                                            "female" = 'Female'
                                            )

graph_ppcs %>%
  group_by(year, civilian_gender) %>%
  mutate(year_count = n()) %>%
  filter(force == 1) %>%
  mutate(force_count = n()) %>%
  summarize(year_count = mean(year_count),
            force_count = mean(force_count)) %>%
  mutate(prop = force_count / year_count) %>%
  ggplot(aes(x = year, y = prop, color = civilian_gender, group = civilian_gender)) +
  geom_point(aes(size = year_count)) + 
  geom_line(lwd = 1.15) + 
  labs(color = "Civilian Gender", 
       size = " Number of Interactions\n Involving Force") +
  xlab("Year") + 
  ylab("Proportion of Use of Force")

```

```{r male-and-female- force #2}
plot2 <- merged_ppcs%>%
  mutate(civilian_gender = recode_factor(civilian_gender, `1` = 'Male',
 `2` = 'Female'))%>%
  group_by(year, civilian_gender)%>%
  mutate(count_year = n())%>%
  filter(force ==1)%>%
  mutate(count_force = n())%>% 
  summarize(count_year = mean(count_year), count_force = mean(count_force))%>% 
  mutate(prop = count_force / count_year) %>%
  ungroup()
  
options(repr.plot.width=6, repr.plot.height=3)

line <- ggplot(plot2, aes(x = year, y= prop , color = civilian_gender, group=civilian_gender)) +
  geom_point(aes(size = count_year))+
  geom_line(lwd = 1) +
  labs(size = "Size"  , color = "Civilian Gender") +
  xlab('Year') +
  ylab('Proportion of use of force')
  
   

line 


```

```{r type-of-force-by-race}
plot01 <- merged_ppcs %>%
  group_by(civilian_race, grab_push)%>%
  filter(grab_push == 1)%>%
  summarize(count_grab = n())
  #dplyr::select(civilian_race, grab_push)

plot02 <- merged_ppcs %>%
  group_by(civilian_race, handcuffed)%>%
  filter(handcuffed == 1)%>%
  summarize(count_handcuffed = n())

plot03 <- merged_ppcs %>%
  group_by(civilian_race, point_gun)%>%
  filter(point_gun == 1)%>%
  summarize(count_point_gun = n())
  

plot04 <- merged_ppcs %>%
  group_by(civilian_race, hit_kick)%>%
  filter(hit_kick == 1)%>%
  summarize(count_hit_kick = n())
  

plot05 <- merged_ppcs %>%
 group_by(civilian_race, pepper_stun)%>%
  filter(pepper_stun == 1)%>%
  summarize(count_pepper_stun = n())
  

forces <- ggplot() + 
  geom_point(data = plot01, aes(y = count_grab, x = civilian_race, size = count_grab, color = "Grab or pushed")) + 
  
  geom_point(data = plot02, aes(y = count_handcuffed, x = civilian_race, size = count_handcuffed, color = "Handcuffed")) +
  
  geom_point(data = plot03, aes(y =count_point_gun, x = civilian_race, size = count_point_gun, color = "Point gun")) +
  
  geom_point(data = plot04, aes(y = count_hit_kick, x = civilian_race, size = count_hit_kick, color = "Hit or Kick"))+
  
  geom_point(data = plot05, aes(y = count_pepper_stun, x = civilian_race, size = count_pepper_stun, color = "Peper or Stun")) +
  labs(color = "Type of force")+
   labs(size = "Size")+
  xlab('Civilian Race') +
  ylab('Size of force used')
  

forces

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
