---
title: "load_1996_ppcs"
author: "Brenda Fried"
date: "7/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries

```{r load-libraries}
library(tidyverse)
library(asciiSetupReader)
library(foreign)
```

#Load Data
```{r load-data}
ppcs_1996 <- read_ascii_setup('06999-0001-Data.txt','06999-0001-Setup.sas')
ppcs_1996 <- ppcs_1996 %>%
  filter(LAST_12_MONTHS_CONTACT_W_POLICE_OFFICER == 'Yes' & CONTACTS_W_POLICE_OFFICER_IN_PERSON == 'Yes')

df<-data.frame(colnames(ppcs_1996)) %>% view
```

#Civilian Race

white if white only and not hispanic
black if black only or black hispanic
hispanic if hispanic and not black
```{r civilian-race}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_race = case_when(
    (RESPONDENT_RACE == "White" & RESPONDENT_ETHNICITY == "Non-Hispanic") ~ "white",
    (RESPONDENT_RACE == "Black" ) ~ "black",
    (RESPONDENT_ETHNICITY == "Hispanic" & RESPONDENT_RACE != "Black") ~ "hispanic",
    TRUE ~ "other"
  ))
```

#Age
```{r civilian-age}
#Age - same as RESPONDENT_AGE column
ppcs_1996 <- ppcs_1996 %>% 
  mutate(civilian_age = NCVS_RESPONDENT_AGE)

```
#Gender
```{r civilian-gender}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_gender = case_when(
    (RESPONDENT_SEX == "Male") ~ 1,
    (RESPONDENT_SEX == "Female") ~ 2
))

```

#Civilian income
Not found
Coded as 1 since 1 is for less than 20k
```{r civilian-income}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_income = NA)

```

#civilian employed
not found so coded as 0
```{r civilian-employed}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_employed = 0)
```
#Population size
not found
coded as 1 since it may be considered as no response
```{r pop-size}
ppcs_1996 <- ppcs_1996 %>%
  mutate(population_size = 1)
```
#Time of Encounter
```{r time-of-encounter}
ppcs_1996 <- ppcs_1996 %>%
  mutate(time_of_encounter = case_when(
    (DAYTIME_OR_NIGHTTIME_WHEN_INCIDENT == "Daytime") ~ 3,
    (DAYTIME_OR_NIGHTTIME_WHEN_INCIDENT == "Nighttime") ~ 6
))
```
#Officer race

Officer Race Black
includes: black, mostly black
```{r officer-race-black}
#officer_race
ppcs_1996 <- ppcs_1996 %>%
  mutate(off_black = case_when(
    (OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'All black' | OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'Mostly black' | OFFICER_WHITE_BLACK_OR_OTHER_RACE == 'Black' ) ~ 1,
    TRUE ~ 0
  ))

```

Officer Race White
includes: white, mostly white

```{r officer-race-white}
ppcs_1996 <- ppcs_1996%>%
  mutate(off_white = case_when(
    (OFFICERS_WHITE_BLACK_OR_OTHER_RACE == "All white" | OFFICERS_WHITE_BLACK_OR_OTHER_RACE == "Mostly white" | OFFICER_WHITE_BLACK_OR_OTHER_RACE == "White" ) ~ 1,
    TRUE ~ 0
  ))

```

Officer Race Other
includes: all of some other race,
mostly some other race

```{r officer-race-other}
ppcs_1996 <- ppcs_1996 %>%
  mutate(off_other = case_when(
    (OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'All of some other race' | OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'Mostly some other race' | OFFICERS_WHITE_BLACK_OR_OTHER_RACE == 'Equally mixed' | OFFICER_WHITE_BLACK_OR_OTHER_RACE == 'Other' ) ~ 1,
    TRUE ~ 0
  ))
```

#Type of incident
Gathered data and then took all the values where the occurence was Once or More than once

 ```{r type-of-incident }
# ppcs_1996 <- ppcs_1996 %>% 
#   gather( key = "incident", value = "occurence", REASON_REPORT_A_CRIME, REASON_ASK_FOR_ASSISTANCE, REASON_PROBLEM_IN_THE_NEIGHBORHOOD,
#          REASON_TRAFFIC_PARKING_VIOLATION, REASON_INVOLVED_IN_TRAFFIC_ACCIDENT, REASON_WITNESS_TO_TRAFFIC_ACCIDENT,
#          REASON_VICTIM_OF_A_CRIME, REASON_WITNESS_TO_A_CRIME, REASON_CRIME_THOUGHT_INVOLVED_IN, REASON_WHAT_YOU_WERE_DOING_IN_THE_AREA,
#          REASON_HAD_WARRANT_FOR_YOUR_ARREST, REASON_CASUAL_ENCOUNTERS, REASON_COMMUNITY_MEETINGS, REASON_SOME_OTHER_REASON,
#          REASON_NO_SPECIFIC_REASON) %>%
#   mutate(type_of_incident = ifelse(occurence == 'Once' | occurence == "More than once", 3, NA)) %>%
#   spread(incident, occurence)
# #ungather/pread so can use again as columns for other variables

#Turns out since every observation is a contact with a police officer they are all 'other stops' and #there should be no NA's
```
```{r type-of-incident}
missing <- c(3,9)
ppcs_1996 <- ppcs_1996 %>%
  mutate(type_of_incident = case_when(
    (REASON_TRAFFIC_PARKING_VIOLATION %in% missing ) ~2,
    (REASON_INVOLVED_IN_TRAFFIC_ACCIDENT %in% missing)~2,
    (REASON_WITNESS_TO_TRAFFIC_ACCIDENT %in% missing)~2,
    TRUE ~ 3
  )) %>% filter(type_of_incident == 2) %>% view
```


#Civilian Behavior
```{r civilian-behavior}
#Civilian behavior
#disobeyed = YOU_DO_INTERFERE_WITH_THE_OFFICER, YOU_DO_BLOCK_OFFICER_S_EXIT_ENTRANCE
#threat = YOU_DO_THREATEN_THE_OFFICER
#force = YOU_DO_ASSAULT_OR_ATTACK_THE_OFFICER
#hit= ??
#complain = ??
#get_away = YOU_DO_ATTEMPT_TO_ESCAPE_HIDE_EVADE
#resist =  YOU_DO_RESIST_PLACED_IN_POL_VEHICLE, YOU_DO_RESIST_BEING_HANDCUFFED
#argue = YOU_DO_ARGUE_WITH_THE_OFFICER
columns <- c(colnames(ppcs_1996)[95:98], colnames(ppcs_1996)[100:103])

is_not_missing <- function(x) {
  ifelse(x == "Not at all" | x == 'missing/ni/ou',0, 1)
}

is_not_missing <- Vectorize(is_not_missing)

ppcs_1996_tmp <- ppcs_1996 %>%
  mutate_at(vars(columns), is_not_missing) %>%
  select(columns)

ppcs_1996 <- ppcs_1996 %>%
  mutate(row_sum = rowSums(ppcs_1996_tmp)) %>%
  mutate(civilian_behavior = case_when(
    (row_sum > 0) ~ 1,
    TRUE ~ 0
  ))

```
#Alternative Outcomes
Civilian searched
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_searched = case_when(
    (DOING_IN_THE_AREA_FRISK_PAT == 'Yes' | CRIME_THOUGHT_INVOLVED_IN_FRISK_PAT == 'Yes') ~ 1,
    TRUE ~ 0
  ))
```
Civilian arrested
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_arrested = case_when(
    (REASON_HAD_WARRANT_FOR_YOUR_ARREST == 'Yes' | HANDCUFF_HAD_WARRANT_FOR_YOUR_ARREST == 'Yes') ~ 1,
    TRUE ~ 0
  ))
```
Civilian guilty of illegal dugs, alcohol, or weapon
```{r}
ppcs_1996 <- ppcs_1996 %>%
  mutate(civilian_guilty_of_illegal = case_when(
    (OTHER_POSSESSION_OF_DRUGS == 'Yes' | OTHER_POSSESSION_OF_FIREARM_WEAPON == 'Yes') ~ 1,
    TRUE ~ 0
  ))
```

##Cleaned Data Frame
```{r}
ppcs_1996_cleaned <- ppcs_1996 %>%
  select(civilian_race, civilian_age, civilian_gender, civilian_income, civilian_employed, population_size, time_of_encounter, off_black, off_white, off_other, type_of_incident, civilian_behavior, civilian_searched, civilian_arrested, civilian_guilty_of_illegal)

View(ppcs_1996_cleaned)
```
#Add year column
```{r}
ppcs_1996_cleaned <- ppcs_1996_cleaned %>%
  mutate(year = 1996)
```

#Save Data
```{r}
save(ppcs_1996_cleaned, file='ppcs_1996.RData')
```

